### 2015-11-29 mcmc.tex
formatting tweaks
model checking pwned

% to-do
% -----
% - Deal with all mentions of ``DFM'' or ``HOGG'' and all ``...''
% - Should we cite the Brooks book or just chapters within it?  I think the latter!
% - Do we need to explain the variance on a sampling-based estimator of an integral?  Right now this is just in an end-note (or will be when I write it).  Should it be in the main text?  Probably!
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - when we say ``below'' also give a Section Number.
% - where to put comments about how to structure code ln_likelihood(pars, data, I) and ln_prior(pars, I)
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

\newcommand{\this}{Using Markov Chain Monte Carlo}
\include{styles/dar}

@@

\begin{document}

\thispagestyle{plain}\sloppy\sloppypar\raggedbottom
\section*{Data analysis recipes:\\ \this\footnotemark}

@@

\paragraph{model checking:}
...Beyond the scope of this document...
Something about cross-validation and posterior predictive checks...
Something about plotting models (drawn from posterior) on data, where appropriate...
How could you possibly know that your sampling really represents your \emph{real} posterior PDF?..

...The Bayes integral / evidence integral?  It is both the biggest strength and biggest weakness that vanilla MCMC methods can't give you this integral.  There are tricks, and then there are MCMC methods.  All that said, you never really want to calculate this integral!

...Posterior predictive checks?

@@

\clearpage
\begin{thebibliography}{}\markright{References}
\bibitem[Brooks \etal(2011)]{mcmchandbook}
  Brooks,~S., Gelman~A., Jones~G., \& Meng,~X.-L., eds.\ 2011,
  \textit{Handbook of Markov Chain Monte Carlo},
  Chapman and Hall~/~CRC Press
  \isbn{978-1-4200-7941-8}.
\bibitem[Foreman-Mackey \etal(2012)]{emcee}
  Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J.\ 2012,
  \project{emcee}: The MCMC Hammer,
  (also \arxiv{1202.3665}).
\bibitem[Hou \etal(2014)]{fengji}
  Hou,~F., Goodman,~J., \& Hogg,~D.~W., 2014,
  The probabilities of orbital-companion models for stellar radial velocity data,
  \arxiv{1401.6128}.
\end{thebibliography}

>>>>

% to-do
% -----
% - Deal with all mentions of ``DFM'' or ``HOGG'' and all ``...''
% - Should we cite the Brooks book or just chapters within it?  I think the latter!
% - Should we make an \acronym{} environment and use it religiously?  I think yes.
% - Do we need to explain the variance on a sampling-based estimator of an integral?  Right now this is just in an end-note (or will be when I write it).  Should it be in the main text?  Probably!
% - mention Gelman--Rubin statistic
% - mention JAGS and STAN in methods section
% - when we say ``below'' also give a Section Number.
% - where to put comments about how to structure code ln_likelihood(pars, data, I) and ln_prior(pars, I)
% - need to mention case where there is an MC integral INSIDE the likelihood function
% - need to mention testing the likelihood function
% - would love to have the Note symbols to be links to the notes and the notes to have links back to the main text...!
% - send to friendlies for comments
%   - Gelman
%   - Lang
%   - Rix
%   - Bovy
%   - Brewer
% - post on arXiv

@@

\newcommand{\chaptertitle}{Using Markov Chain Monte Carlo}
\include{styles/dar}

@@

\begin{document}\sloppy\sloppypar\raggedbottom\thispagestyle{plain}%
%
\section*{Data analysis recipes:\\ \chaptertitle\footnotemark}

@@

\paragraph{model checking:}
\emph{All models are wrong, but some are useful.}
That's not a quote from Box, but it's close.\note{The real quotation
  is a parenthetical sentence ``Remember that all models are wrong;
  the practical question is how wrong do they have to be to not be
  useful.'' (\citealt{box}, p.~74).  It is amazing how rarely this is
  ever quoted correctly; or maybe there is another source for this
  quotation?}
The point (for our purposes) is that if you check your model hard
enough---that is, you take enough data---you will certainly rule it
out.\note{Physicists sometimes forget this, and believe that various
  physical theories, such as quantum electrodynamics or the
  cold-dark-matter cosmological model, are strictly true.  First of
  all, there is no way to know that for sure, and second of all, even
  if they \emph{are} true, they don't precisely explain any
  observation, which is also affected by various kinds of auxilliary
  effects and noises for which there is (and can be, for deep reasons,
  discussed elsewhere by us; CITE) no extremely precise model.  So in
  detail, any model of any specific observation must be wrong, even in
  the (exceedingly unlikely) event that the fundamental model is
  correct.}
``Model checking'' is an enormous subject that goes way beyond the
scope of this \documentname, for which reason we won't say much about
it here except to make two points:

The first---and most important---point about model checking is that
you always learn a huge amount by checking the model.
What this entails is looking at the distribution of residuals or the
quality of the predictions of the model when compared to the data.
There are methods (like the chi-squared statistic and the Bayesian
evidence integral) that check the absolute quality of the model in a
scalar way.
These are usually uninformative, because (as we noted), all models are
wrong, and eventually the data will be good enough to return a bad
statistic here.\note{Sometimes Bayesians like to complain that the
  chi-squared statistic is a measure of the size of your data!  That's
  true when your model is wrong (that is, always).}

Much more informative are methods (like plotting residuals of the data
away from the model in the space of the data) that look at what parts
of the data the model ``explains well'' and what parts of the data the
model ``explains badly''.
These kinds of experiments usually reveal inadequacies of the model in
a rich way, or confirm that certain assumptions are bad.
Cross-validation\note{CITE SOMETHING} is a framework for constructing
valuable tests of this form, as is posterior predictive checks\note{CITE SOMETHING}.
Importantly, you want to find ways to compare your model to your data
\emph{in the space of the data}.
Don't just look at the parameter space of your model!
It is the data that really exist, especially when all models are wrong!

There are model checking methods (like varying the prior and
re-running MCMC) that test the sensitivity of results to assumptions.
Not only are these very valuable for testing the model, they are
essential for writing any paper about your results: The reader always
wants to know the sensitivity of results to assumptions.\note{And be
  sure not to forget that there are (at least) as many assumptions
  encoded in your likelihood function as there are in your prior
  pdfs.}
Once again, it is important to find ways to display the results of
such tests in the space of the data.
All we will say here about these kinds of rich model tests is that you
should do them, and report the results.

The second---and less important---point about model checking is that
the ur-Bayesian thing to be doing is computing a Bayesian evidence
integral.
However, it is usually inadvisable:
The integral is expensive to compute.
Most MCMC methods are cleverly designed to \emph{avoid} computing this
integral; so it usually adds expense to the project.
At the end of the day, it returns only a single, scalar number, with
no guidance about how to use it (and indeed there are only heuristics
to reach for).
If your model is wrong, the richer tests of visualization of residuals
or sensitivity to assumptions are much more likely to lead to insight
about what changes to make or what new directions to explore.

@@

\clearpage
\raggedright
\begin{thebibliography}{}\markright{References}
\bibitem[Box \& Draper(1987)]{box}
  Box,~G.~E.~P. \& Draper,~N.~R.\ 1987,
  \textit{Empirical Model Building and Response Surfaces},
  John Wiley \& Sons, New York, NY
  \isbn{978-0471810339}.
\bibitem[Brooks \etal(2011)]{mcmchandbook}
  Brooks,~S., Gelman~A., Jones~G., \& Meng,~X.-L., eds.\ 2011,
  \textit{Handbook of Markov Chain Monte Carlo},
  Chapman and Hall~/~CRC Press
  \isbn{978-1-4200-7941-8}.
\bibitem[Foreman-Mackey \etal(2012)]{emcee}
  Foreman-Mackey,~D., Hogg,~D.~W., Lang,~D., \& Goodman,~J.\ 2012,
  \project{emcee}: The MCMC Hammer,
  \textit{Pubs.\,Astron.\,Soc.\,Pac.}\ \textbf{125} 306--312
  \doi{10.1086/670067}
  (also \arxiv{1202.3665}).
\bibitem[Hou \etal(2014)]{fengji}
  Hou,~F., Goodman,~J., \& Hogg,~D.~W., 2014,
  The probabilities of orbital-companion models for stellar radial velocity data,
  \arxiv{1401.6128}.
\end{thebibliography}
